Out-of-Sample Comparisons of Overfit Models

### 2014-03-20 paper.tex
Reorganize code (big!)
This commit is a massive code reorganization. There are several reasons
for it:
- R 3.0x came out and my old packages weren't valid any more
- The pgfSweave package is obsolete, and so I've unbundled
  the sweave files that were using it.
- It seemed like an opportunistic time to make a few other
  changes.
Try to resolve problems in pdf fonts
Consolidate header LaTeX into its own file
This commit does some bookkeeping and cleans up the input, usepakage
etc commands. It also puts package loading and macro definitions into
a single LaTeX file.
Styling: tweak spacing and f

\documentclass[11pt]{article}
\usepackage{tikz,setspace}
\usetikzlibrary{calc}
\usetikzlibrary{fit}
\usepackage[round]{natbib}
\usepackage[margin=1in]{geometry}
\onehalfspacing
\input{macros}
\usepackage[charter]{mathdesign}
\usepackage[osf]{sourcecodepro}
% This will make copy and pasting easier, and will prevent the
% abomination that is monospace font ligatures.  Sorry, everyone who
% really likes ligatures.  Don't worry, the Bitstream Charter font
% has very, very few ligatures anyway.
\DisableLigatures{family = tt*}
\urlstyle{same}

\newcommand{\e}{\varepsilon}
\newcommand{\eb}{\mathbf{\e}}
\DeclareMathOperator{\eigen}{\lambda}

\pagestyle{myheadings}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{claim}{Claim}
\newtheorem{cor}{Corollary}
\newtheorem{asmp}{Assumption}
\newtheorem{example}{Example}
\newtheorem{defn}{Definition}

\newcommand{\citepos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}
\newcommand{\Citepos}[1]{\Citeauthor{#1}'s \citeyearpar{#1}}
\newcommand{\clws}{\citeauthor{ClW:06}'s \citeyearpar{ClW:06,ClW:07}}

\usepackage[bitstream]{mathdesign}

% VERSION is generated by a post-commit hook and contains the git
% version info
\IfFileExists{VERSION}{
  \tryinput{VERSION}
  \date{\dinfo\ (commit \hinfo)}
  \markright{\dinfo\ (commit \hinfo)}
}{}

@@

\begin{abstract}\thispagestyle{empty}\noindent
  This paper uses dimension asymptotics to study why overfit linear
  regression models should be compared out-of-sample; we let the
  number of predictors used by the larger model increase with the
  number of observations so that their ratio remains uniformly
  positive.  Our analysis gives a theoretical motivation for using
  out-of-sample (OOS) comparisons: the DMW OOS test allows a forecaster to
  conduct inference about the expected future accuracy of his or her
  models when one or both is overfit.  We show analytically and
  through Monte Carlo that standard full-sample test statistics can
  not test hypotheses about this performance.  Our paper also shows
  that popular test and training sample sizes may give misleading
  results if researchers are concerned about overfit.  We show that
  $P^2/T$ must converge to zero for the DMW test to give valid
  inference about the expected forecast accuracy, otherwise the test
  measures the accuracy of the estimates constructed using only the
  training sample.  In empirical research, $P$ is typically much
  larger than this.  Our simulations indicate that using large values
  of $P$ with the DMW test gives undersized tests with low power, so
  this practice may favor simple benchmark models too much.

@@

\section{Introduction}\label{sec:introduction}
Consider two sequences of length $P$ of prediction errors, the result
of forecasting the same variable with two different estimated models.
Both models are estimated with $R$ observations, collectively called
the {\em estimation window}, and are used to forecast an additional
$P$ observations, called the {\em test sample}.  There are $T$
observations in all, and $R+P=T$.  This paper introduces a new limit
theory for statistics constructed from these prediction errors
designed to approximate the behavior of the statistics when one of the
models is overfit.  In doing so, we provide a theoretical
justification for forecasters to use OOS instead of in-sample
comparisons: the DMW OOS test%
\footnote{In this paper, we will refer to the basic OOS $t$-test
  studied by \citet{DiM:95} and \citet{Wes:96} as the DMW test.} %
allows a forecaster to conduct
inference about the expected future accuracy of his or her models when
one or both is overfit.  We show analytically and through Monte Carlo
that standard full-sample test statistics can not test hypotheses
about this performance.

@@

Although OOS comparisons have been popular in Macroeconomics and
Finance since \citepos{MeR:83} seminal study of exchange rate models,
it has been unclear from a theoretical perspective whether or not the
statistics are useful.  Empirical researchers often cite ``overfit''
or ``instability'' as reasons for using OOS comparisons, as in
\citet{StW:03}, but neither term is precisely defined or formalized.
Compounding this problem, the asymptotic distributions of these
statistics are derived under conditions that rule out either
instability or overfit and allow a researcher to use a conventional
in-sample comparison---a variation of the $F$-test, for example.  As
\citet{InK:04} argue, the statistics themselves are designed to test
hypotheses that can be tested by these in-sample statistics.  For
example, \citet{DiM:95} and \citet{Wes:96} derive the limiting
distributions of many popular OOS test statistics under conditions
that would justify these full-sample tests.  Much
of the subsequent research by \citet{Mcc:00, Mcc:07}, \citet{CCS:01},
\citet{ClM:01,ClM:05}, \citet{CoS:02,CoS:04}, \citet{ClW:06,ClW:07},
\citet{Ana:07}, and others relaxes several of Diebold and Mariano's
and West's assumptions, but maintains the stationarity and dependence
conditions that permit in-sample comparisons \citep[see][for a
review of this literature]{Wes:06}.%
\footnote{Like us, \citet{Ana:07} allows the number of regressors to
  increase with $T$.  But in that paper, the number of regressors
  increases slowly enough that the OLS coefficients are consistent and
  asymptotically normal.} %
\citet{GiW:06} and
\citet{GiR:09, GiR:10} are exceptions.  Instead of focusing on
hypotheses that can be tested by in-sample comparisons, \citet{GiW:06}
derive an OOS test for the null hypothesis that the difference
between two models' OOS forecasting performance is unpredictable, a
martingale difference sequence; \citet{GiR:09} test whether the OOS
forecasting performance of a model suffers from a breakdown relative
to its in-sample performance; and \citet{GiR:10} test whether the
forecasting performance is stable. However, those papers focus on a
particular OOS estimation strategy and do not address why OOS
comparisons might be useful as a general strategy.

@@

In this paper, we study the ``overfit'' possibility and leave
``instability'' to future research. This paper uses dimension
asymptotics to study the behavior of OOS comparisons when at least
one of the models is overfit---the number of its regressors
increases with the number of observations so that their ratio remains
positive.  Although overfit is sometimes used to describe the
situation where a forecaster chooses from many different models,
i.e. \textit{data-mining} or \textit{data-snooping}, we view these as
separate issues.  Procedures that account for the presence of many
models have been, and are continuing to be, developed \citep[see, for
example,][]{Whi:00,Han:05,RoW:05,HHK:10,ClM:12b}, but it is unclear
whether those procedures should themselves use in-sample or OOS
comparisons.  Understanding the difference between in-sample and OOS
comparisons in the context of a simple comparison between two models
is necessary before resolving any new issues that arise with multiple
comparisons. Moreover, the empirical research that motivates
this paper uses \textit{pseudo} OOS comparisons and not
\textit{true} OOS comparisons.  Even if a true OOS comparison
could account for some forms of data-snooping better than
\citepos{Whi:00} BRC or its extensions, in-sample and pseudo
OOS comparisons would both be affected by the data-snooping, a
point also made by \citet{InK:04}.

We focus on linear regression models estimated with a fixed window
for simplicity, but our basic
conclusions should be true for other models as well.  Under this
asymptotic theory, where the number of regressors $k$ increases with $T$
so that $\lim k/T$ is positive and less than one,
the OLS coefficient estimator is no longer
consistent or asymptotically normal \citep{Hub:73} and has positive
variance in the limit.  We show that, even so, the usual OOS average
is asymptotically normal and can consistently estimate the difference
between the models' generalization error, the expected loss in the
future conditional on the on the data available in period
$T$.%
\footnote{See, for example, \citet{HTF:08} for a discussion of
  generalization error.} %
Under these asymptotics, the generalization
error does not converge to the expected performance of the pseudo-true
models, so existing in-sample and OOS comparisons measure different
quantities and should be expected to give different results for
reasons beyond simple size and power comparisons.  Under our limit theory,
the model that is closer to the true DGP in population can forecast
worse.  In such a situation, a standard in-sample comparison would
correctly reject the null hypothesis that the benchmark is true, and
an OOS comparison would correctly fail to reject the null that the
benchmark is more accurate.%
\footnote{In a pair of papers similar to
  ours, \citet{ClM:12,ClM:12b} study in-sample and OOS tests that
  the larger model has nonzero coefficients that are too close to zero
  to expect the model to forecast more accurately.  Like this paper,
  they argue that the larger model can be true but less accurate.
  However, they focus on an aspect of the DGP that makes this
  phenomenon likely, while we focus on the coefficient estimates that
  produce less accurate forecasts.  Moreover, the implications of our
  asymptotic theories are different and their papers do not provide
  reasons to do OOS comparisons.} %
Also note that, in this situation, a model that performs well
in-sample can perform badly out-of-sample even if there are no
structural breaks or other forms of instability.  Researchers
have argued that a breakdown of in-sample forecasting ability
indicates a structural break (see \citealp{BoH:99} and
\citealp{StW:03} among others), but this breakdown can be caused by overfit as
well.

Our results can justify OOS comparisons when researchers want to
choose a model for forecasting.  Although there has been little
emphasis on hypothesis testing in this setting, testing is usually
appropriate: there is usually a familiar benchmark model in place, and
the cost of incorrectly abandoning the benchmark for a less accurate
alternative model is higher than the cost of incorrectly failing to
switch to a more accurate alternative.  We show that the DMW test
lets the forecaster control the probability of the first error, just
as with conventional hypothesis testing.

@@

Popular in-sample tests and model selection criteria, like the Wald
test, AIC, and BIC, do not help forecasters in this setting. We show
that these statistics can no pick out the more accurate model. For
many DGPs they will choose a larger alternative model over a small
benchmark model with probability approaching one, regardless of which
model is more accurate.%
\footnote{Our result holds for a broad class of
  full-sample statistics, but there may be other potential statistics
  that mimic the OOS test and remain valid. Exploring such
  statistics is left for future research.} %
This result holds even though modifications of the $F$-test are valid
under this asymptotic theory, as shown by \citet{BoB:95},
\citet{AkA:00}, \citet{AkP:04}, \citet{Cal:11c}, and \citet{Ana:12},
among others.%
\footnote{Also see \citet{Efr:86,Efr:04} for a discussion
  of naive in-sample loss comparisons.} %
Moreover, under this asymptotic theory, many recent OOS test
statistics, such as those derived by \cite{ClM:01,ClM:05},
\citet{Mcc:07}, and \citet{ClW:06,ClW:07} should have the same
problems as in-sample tests.%
\footnote{Our theoretical results apply
  directly to \citepos{Mcc:07} OOS $t$-test, since it simply
  proposes more liberal critical values for the same test statistic
  that we study.  Since \citet{ClW:06,ClW:07} use a finite length
  estimation window, our asymptotics are incompatible with theirs and
  prevent us from studying their test directly, as well as
  \citepos{GiW:06} and other tests based on \citepos{GiW:06}
  asymptotics.  But \clws\ test can be viewed as a stochastic
  adjustment to the critical values of the usual OOS $t$ test, so
  our conclusions should apply informally as well.  Specifically, we
  show that the DMW test rejects with probability equal to nominal
  size when the estimated benchmark model is expected to be more
  accurate, so more liberal critical values result in overrejection.} %
These tests are also designed to reject the benchmark when the
alternative model is true, and so they may reject too often when the
benchmark is false but more accurate.  Obviously, since the
distribution of these statistics converges to the normal when $P/T \to
0$ (with the number of regressors fixed), these statistics behave like
the DMW test when $P$ is small, but should overreject the benchmark
when $P$ is large. Section~\ref{sec:insample} presents our theoretical
results for full-sample statistics and Section~\ref{sec:mc} presents
Monte Carlo evidence to support these claims.

Finally, this paper also introduces a new method of proof for OOS
statistics.  We use a coupling argument (Berbee's Lemma, 1979) to show
that sequences of OOS loss behave like mixingales when the
underlying series are absolutely regular, even if the forecasts depend
on non-convergent estimators.  Moreover, transformations of these
processes also behave like mixingales when recentered, so 
the arguments used to prove asymptotic
results for Near Epoch Dependent (NED) functions of mixing processes
can be used for these OOS processes with only slight modification.

The rest of the paper proceeds as follows.
Section~\ref{sec:assumptions} introduces our notation and assumptions.
Section~\ref{sec:theory} gives the main theoretical results for the
DMW OOS test, shows that standard in-sample tests reject the
benchmark model too often when it is false but more accurate than the
alternative, and shows that standard model selection methods choose the
larger model too often under similar conditions. Section~\ref{sec:mc}
presents a Monte Carlo study supporting our theoretical results.
Section~\ref{sec:empirics} applies the OOS statistic to
\citepos{GoW:08} dataset for equity premium prediction, and
Section~\ref{sec:conclusion} concludes.  Proofs and supporting results
are listed in the Appendix.

\section{Setup and assumptions}\label{sec:assumptions}

The first part of this section will describe the environment in detail
and set up our models and notation. The second part will list the
assumptions underlying our theoretical results.

\subsection{Notation and forecasting environment}
We assume the following forecasting environment. There are two
competing linear models that give forecasts for the target,
$y_{t+\h}$:
\begin{equation}
y_{t+\h} = x_{1t}'\theta_1 + \e_{1,t+h}, \quad\text{and}\quad
y_{t+\h} = x_{2t}'\theta_2 + \e_{2,t+h};
\end{equation}
$t = 1,\dots,T-h$, $\h$ is the forecast horizon and the
variables $y_t$, $x_{1t}$, and $x_{2t}$ are all known in period $t$.
The coefficients $\theta_1$ and $\theta_2$ minimize the population
Mean Square Error, so
\begin{equation}
  \theta_i = \argmin_\theta \sum_{t=1}^{T-\h} \E (y_{t+\h} - x_{it}'\theta)^2,
\end{equation}
making $\e_{i,t+\h}$ uncorrelated with $x_{i,t}$; $\e_{i,t+\h}$ can exhibit
serial correlation so both of the linear models may be misspecified.
Let $\mathcal{F}_t = \sigma(y_1, x_1, \dots, y_t, x_t)$ be the information
set available in period $t$,
with $x_t$ the vector of all stochastic elements of $x_{1t}$ and $x_{2t}$ after
removing duplicates, and let $\E_t$ and $\var_t$ denote the
conditional mean and variance given $\mathcal{F}_t$.  The first model
uses $K_1$ regressors, and the second uses $K_2$.  Without loss of
generality, assume that $K_1 \leq K_2$. At least one of the models is
overfit, which we represent asymptotically by letting $K_2$ grow with
$T$ quickly enough that $\lim K_2/T$ is positive; $K_1$ may grow with
$T$ as well. Since the models change with $T$, a
stochastic array underlies all of our asymptotic theory, but we
suppress that notation to simplify the presentation.

In the settings we are interested in, a forecaster observes the data
$(y_t,x_t)$ for periods 1 through $T$ and divides the observations
into an estimation sample of the first $R$ observations and a test
sample of the remaining $P$ observations. The forecaster then compares
the models' performance over the test sample, which entails
constructing two sequences of forecasts, (using a fixed-window
estimation strategy)
\begin{equation}
\hat y_{i,t+\h} = x_{it}'\hat{\theta}_{it}, \qquad \text{for } i=1,2;
\ t = R+1,\dots,T-h,
\end{equation}
where
\begin{equation}
  \bh{it} = \Big(\sum_{s=1}^{R-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{R-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = R+1,\dots, T - \h.%
\footnote{It may not be clear why we are using a $t$ index in $\bh{it}$,
    since $\bh{it} = \bh{iR}$ almost surely for all $t \leq T-h$.
    But $\bh{it}$ will be defined for $t > T-\h$
    soon and will not equal $\bh{iR}$ for those values of $t$.} %
\end{equation}

The models are then compared by their forecast errors over the test
sample.  There are many statistics that have been considered in the
literature, but we focus on perhaps the most natural, the DMW \oost\
test \citep{DiM:95,Wes:96}.%
\footnote{The core insights of our paper apply to other OOS statistics
  as well.} %
This statistic is based on
the difference in the models' loss over the test sample, $\oosA$,
defined as
\begin{equation}
  \oosA \equiv P^{-1} \oosSum{t}{1} D_t
\end{equation}
where
\begin{equation}\label{eq:16}
  D_t = L(y_{t+\h} - x_{1t}'\bh{1t}) - L(y_{t+\h} - x_{2t}'\bh{2t}),
\end{equation}
and $L$ is a known loss function. The \oost\ test is defined as
$\sqrt{P} \oosA / \sh$, where $\sh^2$ is an estimator of the
asymptotic variance of $\oosA$. (Possibly a Heteroskedasticity- and
Autocorrelation-Consistent, or HAC, estimator.)

Most theoretical research, such as \citet{DiM:95}, \citet{Wes:96},
and \cite{Mcc:07}, has focused on using the \oost\ statistic to test
hypotheses about the pseudo-true values $\theta_1$ and $\theta_2$.  In
particular, that research focuses on testing the null hypothesis that
\begin{equation}\label{eq:21}
  \E L(y_{t+\h} - x_{1t}'\theta_{1}) - \E L(y_{t+\h} - x_{2t}'\theta_{2}) = 0.
\end{equation}
But the population quantities in \eqref{eq:21} do not determine which
model is more accurate in finite samples and, under our asymptotics,
in the limit when models are allowed to be overfit. The models'
accuracy will also depend on the specific estimates of $\theta_1$ and
$\theta_2$ used to produce the forecasts. In this setting, where the
forecaster will use one of the models to make a number of predictions
in the future---call the number of predictions $Q$---the quantity of
interest becomes
\begin{equation}
  \E_T \oosB \equiv Q^{-1} \sum_{t=T+1}^{T+Q} \E_T D_t,
\end{equation}
where $D_t$ is defined in Equation~\eqref{eq:16} but here uses the
full-sample estimates of the models' parameters,
\begin{equation}
  \bh{it} = \Big(\sum_{s=1}^{T-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{T-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = T+1,\dots, T + Q.
\end{equation}
If $\E_T \oosB$ is positive, the second model is expected to forecast
better than the first over the next $Q$ periods, and if $\E_T \oosB$
is negative then the first model is better. We use a conditional
expectation because the coefficient estimates in $\oosB$ are
stochastic but known in period $T$, and their values will determine
the performance of the two models.

>>>>

\documentclass[12pt]{article}
\input{setup}

@@

\begin{abstract}\noindent
  This paper uses dimension asymptotics to study why overfit linear
  regression models should be compared out-of-sample; we let the
  number of predictors used by the larger model increase with the
  number of observations so that their ratio remains uniformly
  positive.  Our analysis gives a theoretical motivation for using
  out-of-sample (OOS) comparisons: the DMW OOS test allows a forecaster to
  conduct inference about the expected future accuracy of his or her
  models when one or both is overfit.  We show analytically and
  through Monte Carlo that standard full-sample test statistics can
  not test hypotheses about this performance.  Our paper also shows
  that popular test and training sample sizes may give misleading
  results if researchers are concerned about overfit.  We show that
  $P^2/T$ must converge to zero for the DMW test to give valid
  inference about the expected forecast accuracy, otherwise the test
  measures the accuracy of the estimates constructed using only the
  training sample.  In empirical research, $P$ is typically much
  larger than this.  Our simulations indicate that using large values
  of $P$ with the DMW test gives undersized tests with low power, so
  this practice may favor simple benchmark models too much.

@@

\section{Introduction}
\label{sec:introduction}

Consider two sequences of length $P$ of prediction errors, the result
of forecasting the same variable with two different estimated models.
Both models are estimated with $R$ observations, collectively called
the {\em estimation window}, and are used to forecast an additional
$P$ observations, called the {\em test sample}.  There are $T$
observations in all, and $R+P=T$.  This paper introduces a new limit
theory for statistics constructed from these prediction errors
designed to approximate the behavior of the statistics when one of the
models is overfit.  In doing so, we provide a theoretical
justification for forecasters to use OOS instead of in-sample
comparisons: the DMW OOS test%
\footnote{In this paper, we will refer to the basic OOS $t$-test
  studied by \citet{DiM:95} and \citet{Wes:96} as the DMW test.} %
allows a forecaster to conduct
inference about the expected future accuracy of his or her models when
one or both is overfit.  We show analytically and through Monte Carlo
that standard full-sample test statistics can not test hypotheses
about this performance.

@@

Although OOS comparisons have been popular in Macroeconomics and
Finance since \citepos{MeR:83} seminal study of exchange rate models,
it has been unclear from a theoretical perspective whether or not the
statistics are useful.  Empirical researchers often cite ``overfit''
or ``instability'' as reasons for using OOS comparisons, as in
\citet{StW:03}, but neither term is precisely defined or formalized.
Compounding this problem, the asymptotic distributions of these
statistics are derived under conditions that rule out either
instability or overfit and allow a researcher to use a conventional
in-sample comparison---a variation of the $F$-test, for example.  As
\citet{InK:04} argue, the statistics themselves are designed to test
hypotheses that can be tested by these in-sample statistics.  For
example, \citet{DiM:95} and \citet{Wes:96} derive the limiting
distributions of many popular OOS test statistics under conditions
that would justify these full-sample tests.  Much
of the subsequent research by \citet{Mcc:00, Mcc:07}, \citet{CCS:01},
\citet{ClM:01,ClM:05}, \citet{CoS:02,CoS:04}, \citet{ClW:06,ClW:07},
\citet{Ana:07}, and others relaxes several of Diebold and Mariano's
and West's assumptions, but maintains the stationarity and dependence
conditions that permit in-sample comparisons \citep[see][for a
review of this literature]{Wes:06}.%
\footnote{Like us, \citet{Ana:07} allows the number of regressors to
  increase with $T$.  But in that paper, the number of regressors
  increases slowly enough that the OLS coefficients are consistent and
  asymptotically normal.} %
\citet{GiW:06} and
\citet{GiR:09, GiR:10} are exceptions.  Instead of focusing on
hypotheses that can be tested by in-sample comparisons, \citet{GiW:06}
derive an OOS test for the null hypothesis that the difference
between two models' OOS forecasting performance is unpredictable, a
martingale difference sequence (MDS); \citet{GiR:09} test whether the OOS
forecasting performance of a model suffers from a breakdown relative
to its in-sample performance; and \citet{GiR:10} test whether the
forecasting performance is stable. However, those papers focus on a
particular OOS estimation strategy and do not address why OOS
comparisons might be useful as a general strategy.

@@

In this paper, we study the ``overfit'' possibility and leave
``instability'' to future research. This paper uses dimension
asymptotics to study the behavior of OOS comparisons when at least
one of the models is overfit---the number of its regressors
increases with the number of observations so that their ratio remains
positive.  Although overfit is sometimes used to describe the
situation where a forecaster chooses from many different models,
i.e. \emph{data-mining} or \emph{data-snooping}, we view these as
separate issues.  Procedures that account for the presence of many
models have been, and are continuing to be, developed \citep[see, for
example,][]{Whi:00,Han:05,RoW:05,HHK:10,ClM:12b}, but it is unclear
whether those procedures should themselves use in-sample or OOS
comparisons.  Understanding the difference between in-sample and OOS
comparisons in the context of a simple comparison between two models
is necessary before resolving any new issues that arise with multiple
comparisons. Moreover, the empirical research that motivates
this paper uses \emph{pseudo} OOS comparisons and not
true OOS comparisons.  Even if a true OOS comparison
could account for some forms of data-snooping better than
\citepos{Whi:00} BRC or its extensions, in-sample and pseudo
OOS comparisons would both be affected by the data-snooping, a
point also made by \citet{InK:04}.

We focus on linear regression models estimated with a fixed window
for simplicity, but our basic conclusions should be true for other
models and estimation strategies as well.  Under this
asymptotic theory, where the number of regressors $k$ increases with $T$
so that $\lim k/T$ is positive and less than one,
the OLS coefficient estimator is no longer
consistent or asymptotically normal \citep{Hub:73} and has positive
variance in the limit.  We show that, even so, the usual OOS average
is asymptotically normal and can consistently estimate the difference
between the models' generalization error, the expected loss in the
future conditional on the on the data available in period
$T$.%
\footnote{See, for example, \citet{HTF:08} for a discussion of
  generalization error.} %
Under these asymptotics, the generalization
error does not converge to the expected performance of the pseudotrue
models, so existing in-sample and OOS comparisons measure different
quantities and should be expected to give different results for
reasons beyond simple size and power comparisons.  Under our limit theory,
the model that is closer to the true DGP in population can forecast
worse.  In such a situation, a standard in-sample comparison would
correctly reject the null hypothesis that the benchmark is true, and
an OOS comparison would correctly fail to reject the null that the
benchmark is more accurate.%
\footnote{In a pair of papers similar to
  ours, \citet{ClM:12,ClM:12b} study in-sample and OOS tests that
  the larger model has nonzero coefficients that are too close to zero
  to expect the model to forecast more accurately.  Like this paper,
  they argue that the larger model can be true but less accurate.
  However, they focus on an aspect of the DGP that makes this
  phenomenon likely, while we focus on the coefficient estimates that
  produce less accurate forecasts.  Moreover, the implications of our
  asymptotic theories are different and their papers do not provide
  reasons to do OOS comparisons.} %
Also note that, in this situation, a model that performs well
in-sample can perform badly out-of-sample even if there are no
structural breaks or other forms of instability.  Researchers
have argued that a breakdown of in-sample forecasting ability
indicates a structural break (see \citealp{BoH:99}, and
\citealp{StW:03}, among others), but this breakdown can be caused by
overfit as well.

These theoretical results partially justify OOS comparisons when
researchers want to
choose a model for forecasting.  Although there has been little
emphasis on hypothesis testing in this setting, testing is usually
appropriate: there is usually a familiar benchmark model in place, and
the cost of incorrectly abandoning the benchmark for a less accurate
alternative model is higher than the cost of incorrectly failing to
switch to a more accurate alternative.  We show that the DMW test
lets the forecaster control the probability of the first error, just
as with conventional hypothesis testing.

@@

Popular in-sample tests and model selection criteria, like the Wald
test and the AIC, do not help forecasters in this setting. We show
that these statistics do not select the more accurate model when
choosing between overfit models. For many DGPs the Wald test and the
AIC will choose a larger alternative model over a small benchmark
model with probability approaching one, regardless of which model is
more accurate, while the BIC chooses the benchmark model with
probability approaching one.%
\footnote{Our result holds for a broad class of
  full-sample statistics, but there may be other potential statistics
  that mimic the OOS test and remain valid. Exploring such
  statistics is left for future research.} %
This result holds even though modifications of the $F$-test are valid
under this asymptotic theory, as shown by \citet{BoB:95},
\citet{AkA:00}, \citet{AkP:04}, \citet{Cal:11c}, and \citet{Ana:12},
among others.%
\footnote{Also see \citet{Efr:86,Efr:04} for a discussion
  of naive in-sample loss comparisons.} %
Moreover, under this asymptotic theory, many recent OOS test
statistics, such as those derived by \cite{ClM:01,ClM:05},
\citet{Mcc:07}, and \citet{ClW:06,ClW:07} should have the same
problems as in-sample tests.%
\footnote{Our theoretical results apply
  directly to \citepos{Mcc:07} OOS $t$-test, since it simply
  proposes more liberal critical values for the same test statistic
  that we study.  Since \citet{ClW:06,ClW:07} use a finite length
  estimation window, our asymptotics are incompatible with theirs and
  prevent us from studying their test directly, as well as
  \citepos{GiW:06} and other tests based on \citepos{GiW:06}
  asymptotics.  But \clws\ test can be viewed as a stochastic
  adjustment to the critical values of the usual OOS $t$ test, so
  our conclusions should apply informally as well.  Specifically, we
  show that the DMW test rejects with probability equal to nominal
  size when the estimated benchmark model is expected to be more
  accurate, so more liberal critical values result in overrejection.} %
These tests are also designed to reject the benchmark when the
alternative model is true, and so they may reject too often when the
benchmark is false but more accurate.  Obviously, since the
distribution of these statistics converges to the normal when $P/T \to
0$ (with the number of regressors fixed), these statistics behave like
the DMW test when $P$ is small, but should overreject the benchmark
when $P$ is large. Section~\ref{sec:insample} presents our theoretical
results for full-sample statistics and Section~\ref{sec:mc} presents
Monte Carlo evidence to support these claims.

Finally, this paper introduces a new method of proof for OOS
statistics.  We use a coupling argument (Berbee's Lemma, 1979) to show
that sequences of OOS loss behave like mixingales when the
underlying series are absolutely regular, even if the forecasts depend
on non-convergent estimators.  Moreover, transformations of these
processes also behave like mixingales when recentered, so 
the arguments used to prove asymptotic
results for Near Epoch Dependent (NED) functions of mixing processes
can be used for these OOS processes with only slight modification.

The rest of the paper proceeds as follows.
Section~\ref{sec:assumptions} introduces our notation and assumptions.
Section~\ref{sec:theory} gives the main theoretical results for the
DMW OOS test, shows that standard in-sample tests reject the
benchmark model too often when it is false but more accurate than the
alternative, and shows that standard model selection methods can run
into similar problems. Section~\ref{sec:mc}
presents a Monte Carlo study supporting our theoretical results.
Section~\ref{sec:empirics} applies the OOS statistic to
\citepos{GoW:08} dataset for equity premium prediction, and
Section~\ref{sec:conclusion} concludes.  Proofs and supporting results
are listed in the Appendix.

\section{Setup and assumptions}
\label{sec:assumptions}

The first part of this section will describe the environment in detail
and set up our models and notation. The second part lists the
assumptions underlying our theoretical results.

\subsection{Notation and forecasting environment}

We assume the following forecasting environment. There are two
competing linear models that give forecasts for the target,
$y_{t+\h}$:
\begin{equation}
y_{t+\h} = x_{1t}'\theta_1 + \e_{1,t+h}, \quad\text{and}\quad
y_{t+\h} = x_{2t}'\theta_2 + \e_{2,t+h};
\end{equation}
$t = 1,\dots,T-h$, $\h$ is the forecast horizon and the
variables $y_t$, $x_{1t}$, and $x_{2t}$ are all known in period $t$.
The coefficients $\theta_1$ and $\theta_2$ minimize the population
Mean Square Error, so
\begin{equation}
  \theta_i = \argmin_\theta \sum_{t=1}^{T-\h} \E (y_{t+\h} - x_{it}'\theta)^2,
\end{equation}
making $\e_{i,t+\h}$ uncorrelated with $x_{i,t}$; $\e_{i,t+\h}$ can exhibit
serial correlation so both of the linear models may be misspecified.
Let $\mathcal{F}_t = \sigma(y_1, x_1, \dots, y_t, x_t)$ be the information
set available in period $t$,
with $x_t$ the vector of all stochastic elements of $x_{1t}$ and $x_{2t}$ after
removing duplicates, and let $\E_t$ and $\var_t$ denote the
conditional mean and variance given $\mathcal{F}_t$.  The first model
uses $K_1$ regressors, and the second uses $K_2$.  Without loss of
generality, assume that $K_1 \leq K_2$. At least one of the models is
overfit, which we represent asymptotically by letting $K_2$ grow with
$T$ quickly enough that $\lim K_2/T$ is positive; $K_1$ may grow with
$T$ as well. Since the models change with $T$, a
stochastic array underlies all of our asymptotic theory, but we
suppress that notation to simplify the presentation.

In the settings we are interested in, a forecaster observes the data
$(y_t,x_t)$ for periods 1 through $T$ and divides the observations
into an estimation sample of the first $R$ observations and a test
sample of the remaining $P$ observations. The forecaster then compares
the models' performance over the test sample, which entails
constructing two sequences of forecasts, (using a fixed-window
estimation strategy)
\begin{equation}
\hat y_{i,t+\h} = x_{it}'\hat{\theta}_{it}, \qquad \text{for } i=1,2;
\ t = R+1,\dots,T-h,
\end{equation}
where
\begin{equation}
  \bh{it} = \Big(\sum_{s=1}^{R-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{R-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = R+1,\dots, T - \h.%
\footnote{It may not be clear why we are using the index $t$ in $\bh{it}$,
    since $\bh{it} = \bh{iR}$ almost surely for all $t \leq T-h$.
    But $\bh{it}$ will be defined for $t > T-\h$
    soon and will not equal $\bh{iR}$ for those values of $t$.} %
\end{equation}

The models are then compared by their forecast performance over the test
sample. There are many statistics that have been considered in the
literature, but we focus on perhaps the most natural, the DMW \oost\
test \citep{DiM:95,Wes:96}.%
\footnote{The core insights of our paper apply to other OOS statistics
  as well.} %
This statistic is based on
the difference in the models' loss over the test sample, $\oosA$,
defined as
\begin{equation}
  \oosA \equiv P^{-1} \oosSum{t}{1} D_t
\end{equation}
where
\begin{equation}\label{eq:16}
  D_t = L(y_{t+\h} - x_{1t}'\bh{1t}) - L(y_{t+\h} - x_{2t}'\bh{2t}),
\end{equation}
and $L$ is a known loss function. The \oost\ test is defined as
$\sqrt{P} \oosA / \sh$, where $\sh^2$ is an estimator of the
asymptotic variance of $\oosA$. (Possibly a Heteroskedasticity- and
Autocorrelation-Consistent, or HAC, estimator.)

Statistics like $\oost$ have a long history in empirical economics
because they capture an intuitive idea of modeling: that a good model
should be able to forecast well when given new data and not only when
used to fit the same data that were used to estimate it. \cite{MeR:83}
use such a statistic to study exchange rate models and find that none
of the models that existed at the time of the study outperformed a
random walk. This finding has been remarkably durable and has spawned
an enormous literature; see \cite{Mar:95}, \cite{KiT:03},
\cite{CCP:05}, \cite{EnW:05}, \cite{Ros:05}, and \cite{BWB:10}, among
many others.

Most theoretical research on these statistics, such as \citet{DiM:95}, \citet{Wes:96},
and \cite{Mcc:07}, has focused on using the \oost\ statistic to test
hypotheses about the pseudotrue values $\theta_1$ and $\theta_2$.  In
particular, that research focuses on testing the null hypothesis that
\begin{equation}\label{eq:21}
  \E L(y_{t+\h} - x_{1t}'\theta_{1}) - \E L(y_{t+\h} - x_{2t}'\theta_{2}) = 0.
\end{equation}
But the population quantities in \eqref{eq:21} do not determine which
model is more accurate in practice. The models'
accuracy will also depend on the specific estimates of $\theta_1$ and
$\theta_2$ used to produce the forecasts. In this setting, where the
forecaster will use one of the models to make a number of predictions
in the future---call the number of predictions $Q$---the quantity of
interest becomes
\begin{equation}
  \E_T \oosB = Q^{-1} \sum_{t=T+1}^{T+Q} \E_T D_t,
\end{equation}
where $D_t$ is defined in Equation~\eqref{eq:16} but here uses the
full-sample estimates of the models' parameters,
\begin{equation}
  \bh{it} = \Big(\sum_{s=1}^{T-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{T-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = T+1,\dots, T + Q.
\end{equation}
If $\E_T \oosB$ is positive, the second model is expected to forecast
better than the first over the next $Q$ periods, and if $\E_T \oosB$
is negative then the first model is better. We use a conditional
expectation because the coefficient estimates in $\oosB$ are
stochastic but known in period $T$, and their values will determine
the performance of the two models.
